From Freebase to Wikidata: The Great Migration

### 2015-10-15 paper.tex
Updated intro

\begin{abstract}
Collaborative knowledge bases that make their data freely available in a~machine-readable form
are central for the data strategy of many projects and organizations.
One such collaborative knowledge base is Freebase, initially launched by Metaweb in 2007,
who were then acquired by Google in 2010.
Another example is Wikidata, a~collaborative knowledge base developed by
Wikimedia Deutschland since 2012 and operated by the Wikimedia Foundation.
Due to the success of Wikidata, Google decided 2015 to shut down Freebase and
help the community with the transfer of Freebase's content to Wikidata.
In this paper, we describe the ongoing transfer efforts and data mapping challenges,
and provide an analysis of the effort so far.
We have created the \emph{Primary Sources Tool}
that we have released as open source to facilitate the transfer of this dataset,
but also future data migrations.
Throughout the migration, we have gained insights into both Wikidata and Freebase,
and share and discuss detailed statistics on both knowledge bases. 
\end{abstract}

\category{H.3.5}{Information Storage and Retrieval}{Online Information Systems}
\category{H.5.3}{Information Interfaces}{Group and Organization Interfaces?Web-based interactions}
\category{I.2.4}{Artifical Intelligence}{Knowledge Representation}
\category{K.4.3}{Computers and Society}{Organizational Impacts Computer-supported collaborative work}

@@

Moving data between two knowledge bases that do not share a~similar design
is usually a~problematic task and requires the careful mapping between their structures.
In this paper, we describe how we facilitated the migration of
Freebase\footnote{Freebase: \url{https://www.freebase.com}} content to
Wikidata.\footnote{Wikidata: \url{https://www.wikidata.org}}
That migration was no exception to this rule, and had a~number of \emph{structural} challenges.
Even more challenging was the \emph{cultural} difference between the two involved communities.
The Freebase and Wikidata communities have a~very different background,
subtly different goals and understandings of their tasks,
and different requirements regarding their data.
\todo{Expand}

The remainder of this paper is structured as follows.
After an introduction to the two collaborative knowledge bases
Freebase and Wikidata in \autoref{sec:background},
we describe our methodology and the metrics used to measure the migration
in \autoref{sec:challenges-of-the-migration}.
In order to support the migration, we have developed a~set of open source tools
that we present in \autoref{sec:primary-sources-tool}.
We then present the results of the migration
and discuss these statistics in \autoref{sec:statistics-of-the-migration}.
In \autoref{sec:related-work} we discuss the relation to other work.
The paper terminates with an outlook at proposed next steps
and a~conclusion in \autoref{sec:future-work-and-conclusion}.

@@

Freebase is an open and
collaborative knowledge base created in 2007 by Metaweb and 
acquired in 2010 by Google~\cite{freebase}.
It was used as the open core of the Google Knowledge Graph,
and has found many use cases outside of Google.
Due to the success of Wikidata,
Google announced in 2014 to close Freebase and offer most of its content to the
Wikidata community~\cite{freebase-end}.

Freebase relies on the notions of \emph{topics}, \emph{facts}, \emph{types}, and \emph{properties}.
Each Freebase topic has a~stable identifier called \emph{``mid''} (for Metaweb ID),
one or more types, and uses properties from these types in order to provide facts.
For example, the Freebase topic for Barack Obama has the mid \texttt{/m/02mjmr}
and the type \texttt{/government/us\_president} (among others) that allows the entity to have
a fact with the property \texttt{/government/us\_president/presidency\_number}
and the literal integer ``44'' as the value.

@@

Freebase uses Compound Value Types (CVTs, see \autoref{fig:cvt-obama}) to represent
n-ary relations with $n>2$, e.g. values like geographic coordinates,
political positions held with a start and end date, or actors playing a specific character in a movie.
CVT values are much like topics, \emph{i.e.} they have a~mid and can have types
(usually only the compound value type).

The content of Freebase has been imported from various sources like Wikipedia
or the Internet Movie Database IMDB,
in consequence Google does not own the copyright of some parts of the content of the knowledge base
like images or long entity descriptions extracted from Wikipedia.
As of the closing of Freebase on March~31, 2015,
it counted more than 3~billion facts about almost 50M~entities.

@@

Wikidata\footnote{Wikidata: \url{https://www.wikidata.org}}
is a~collaborative knowledge base
launched in October 2012 and hosted by the Wikimedia Foundation.
Its community has been growing quickly since the beginning, and as of mid 2015,
the community comprises about 6,000 active contributors.

Wikidata's data model relies on the notions of \emph{item} and \emph{statement}.
An item represent a~concept, has a~stable identifier and may have some labels,
descriptions, and aliases in multiple languages, links to pages about this concept
in the others Wikimedia projects, like Wikipedia, and statements.
Contrary to Freebase, Wikidata encodes with its statements not true facts,
but \emph{claims} from different sources.

A~statement is composed of one claim and zero or more references for this claim.
The claim itself is composed of one main property--value couple that encodes
the main (claimed) fact like ``population is 8,173,900'' and optional qualifiers
to add information about it like ``as of June 2012''.
\autoref{fig:statement} illustrates the used nomenclature with an example.

The content of Wikidata is freely available under a~Creative Commons~0 (CC0) license.
As of mid 2015, Wikidata counted about 66M~statements on 14M~entities.
For more background information about Wikidata, see~\cite{vrandevcic2014wikidata}
by Vrandečić and Krötzsch.
For a~comparison between Freebase and Wikidata,
we refer the reader to~\cite{farbercomparative} by Färber \emph{et~al.}

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/Wikidata_statement.pdf}
\caption{Wikidata nomenclature illustrated with an example (Source:
	\url{https://commons.wikimedia.org/wiki/File:Wikidata_statement.svg}).}
\label{fig:statement}
\end{figure}

@@

The first challenge concerns the licenses under which the datasets are published.
Wikidata is published under a~Creative Commons~0 (CC0~1.0) license
that effectively puts the data in the public domain,
whereas Freebase is published under a~Creative Commons Attribution (CC~BY~2.5) license
and contains data imported from external projects not owned by Google,
like images or descriptions extracted from Wikipedia.
In a~first step, we filtered the Freebase dump for this kind of content
before being able to create a~dump that we could relicense under CC0~1.0.
This step reduced the number of Freebase facts that could be republished by about 42M~facts
from 3~billion facts in the original corpus.

>>>>

\begin{abstract}
Collaborative knowledge bases that make their data freely available in a~machine-readable form
are central for the data strategy of many projects and organizations.
The two major collaborative knowledge bases are Google's Freebase and Wikimedia's Wikidata.
Due to the success of Wikidata, Google decided in 2015 to offer the content of Freebase
to the Wikidata community.
In this paper, we describe the ongoing transfer efforts and data mapping challenges,
and provide an analysis of the effort so far.
We describe the \emph{Primary Sources Tool}, which aims to facilitate
this and future data migrations.
Throughout the migration, we have gained insights into both Wikidata and Freebase,
and share and discuss detailed statistics on both knowledge bases. 
\end{abstract}

\category{H.3.5}{Information Storage and Retrieval}{Online Information Systems}
\category{H.5.3}{Information Interfaces}{Group and Organization Interfaces?Web-based interactions}
\category{I.2.4}{Artificial Intelligence}{Knowledge Representation}
\category{K.4.3}{Computers and Society}{Organizational Impacts Computer-supported collaborative work}

@@

Large web-based knowledge bases that make their data available
under free licenses in a~machine-readable form
have become central for the data strategy of many projects and organizations.
They find applications in areas as diverse as web search,
natural language annotation, or translation.
\todo{add references to papers with examples}

One such collaborative knowledge base is Freebase,
initially launched by Metaweb in 2007 and acquired by Google in 2010.
Another example is Wikidata, a~collaborative knowledge base developed by
Wikimedia Deutschland since 2012 and operated by the Wikimedia Foundation.
Due to the success of Wikidata, Google announced in 2014 to shut down Freebase and
help the community with the transfer of Freebase's content to Wikidata.

Moving data between two knowledge bases that do not share a~similar design
is usually a~problematic task and requires the careful mapping between their structures.
The migration from Freebase to Wikidata was no exception to this rule:
we encountered a~number of \emph{structural} challenges.
Even more challenging was the \emph{cultural} difference between the two involved communities.
The Freebase and Wikidata communities have a~very different background,
subtly different goals and understandings of their tasks,
and different requirements regarding their data.

In this paper we describe how we support the Wikidata community with
the migration of Freebase content to Wikidata.
We created the \emph{Primary Sources Tool},
released as open source to facilitate the transfer of the Freebase dataset.
The \emph{Primary Sources Tool} is also developed with an eye on
future data migrations from other source.
We created and published mappings of the Freebase dataset.
Throughout the migration, we have gained insights into both Wikidata and Freebase,
and share and discuss detailed statistics on both knowledge bases.

The paper is structured as follows.
First, we introduce Freebase and Wikidata in \autoref{sec:background}.
Then we describe our methodology and the metrics used to measure the migration
in \autoref{sec:challenges-of-the-migration}.
In order to support the migration, we have developed a~set of open source tools
that we present in \autoref{sec:primary-sources-tool}.
We then present the results of the migration
and discuss these statistics in \autoref{sec:statistics-of-the-migration}.
Before we close, we discuss related work in \autoref{sec:related-work},
followed by an outlook at proposed next steps
and a~conclusion in \autoref{sec:future-work-and-conclusion}.

@@

Freebase\footnote{Freebase: \url{https://www.freebase.com}} is an open and
collaborative knowledge base created in 2007 by Metaweb and 
acquired in 2010 by Google~\cite{freebase}.
It was used as the open core of the Google Knowledge Graph,
and has found many use cases outside of Google.
Due to the success of Wikidata,
Google announced in 2014 to close Freebase and offer most of its content to the
Wikidata community~\cite{freebase-end}.

Freebase is build on the notions of \emph{objects}, \emph{facts}, \emph{types}, and \emph{properties}.
Each Freebase object has a~stable identifier called a \emph{``mid''} (for Metaweb ID),
one or more types, and uses properties from these types in order to provide facts.
For example, the Freebase object for Barack Obama has the mid \texttt{/m/02mjmr}
and the type \texttt{/government/us\_president} (among others) that allows the entity to have
a fact with the property \texttt{/government/us\_president/presidency\_number}
and the literal integer ``44'' as the value.

@@

Freebase uses Compound Value Types (CVTs, see \autoref{fig:cvt-obama}) to represent
n-ary relations with $n>2$, e.g. values like geographic coordinates,
political positions held with a start and end date, or actors playing a character in a movie.
CVT values are just objects, \emph{i.e.} they have a~mid and can have types
(although they usually have only the compound value type itself).
Most non-CVT objects are called \emph{topics} in order to discern them from CVTs.

The content of Freebase has been partially imported from various sources like Wikipedia
or the license-compatible part of MusicBrainz.
Over the years, the Freebase community and Google have maintained the knowledge base.
When Freebase was turned read-only on March~31, 2015,
it counted more than 3~billion facts about almost 50M~entities.

The data from Freebase is published as an N-Triples dump in RDF~\cite{rdf} under
the Creative Commons CC-BY license.\footnote{\url{https://developers.google.com/freebase/data}}

@@

Wikidata\footnote{Wikidata: \url{https://www.wikidata.org}}
is a~collaborative knowledge base
launched in October 2012 and hosted by the Wikimedia Foundation~\cite{vrandevcic2014wikidata}.
Its community has been growing quickly, and as of mid 2015,
the community comprises about 6,000 active contributors~\cite{wikidatastats}.

Wikidata's data model relies on the notions of \emph{item} and \emph{statement}.
An item represents an~entity, has a~stable identifier and may have labels,
descriptions, and aliases in multiple languages, links to pages about this concept
in the others Wikimedia projects, most prominently Wikipedia, and statements.
Contrary to Freebase, Wikidata statements do not aim to encode true facts,
but \emph{claims} from different sources which can also contradict each other.

A~statement is composed of a claim and zero or more references for this claim.
The claim itself is composed of one main property--value couple that encodes
the main (claimed) fact like ``population is 8,173,900'' and optional qualifiers
to add information about it like ``as of June 2012''.
\autoref{fig:statement} illustrates the terminology with an example.

The content of Wikidata is available under the Creative Commons CC0 license.
As of mid 2015, Wikidata counted about 66M~statements on 14M~entities. \todo{update}

A~comparison between Freebase, Wikidata, and other free and open knowledge bases
(DBpedia, YAGO, and OpenCyc) is available in~\cite{farbercomparative}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/Wikidata_statement.pdf}
\caption{Wikidata terminology illustrated with an example (Source:
	\url{https://commons.wikimedia.org/wiki/File:Wikidata_statement.svg}).}
\label{fig:statement}
\end{figure}

@@

The first challenge concerns the licenses under which the datasets are published.
Wikidata is published under a~Creative Commons~0 (CC0~1.0) license
that effectively puts the data in the public domain,
whereas Freebase is published under a~Creative Commons Attribution (CC~BY~2.5) license
and contains data imported from external projects not owned by Google,
like images or descriptions extracted from Wikipedia.
In consequence Google does not own the copyright of some parts of the content of the knowledge base,
such as images or long entity descriptions extracted from Wikipedia.
In a~first step, we filtered the Freebase dump for this kind of content
before being able to create a~dump that we could relicense under CC0~1.0.
This step reduced the number of Freebase facts that could be republished by about 42M~facts
from 3~billion facts in the original corpus.
